APPM 3310 Project\\\Large{Static Analysis of Truss Systems}

### 2016-04-25 paper.tex
Add merge section.

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumitem}

@@

Most of the following formulation follows directly from
\parencite{felippa2004}.

@@

The direct stiffness method involves 8 steps.  The first two are purely
conceptual and thus are never really performed and the last one is a post
processing step performed after a solution is found.  These eight steps are
outlined below (taken from Figure 2.5 of \cite{felippa2004}).
\begin{enumerate}[noitemsep]
    \item Disconnection
    \item Localization
    \item Element Formation
    \item Globalization
    \item Merge
    \item Application of Boundary Conditions
    \item Solution
    \item Recovery of Derived Quantities
\end{enumerate}

@@

Before the global stiffness matrix can be formed
Eq.~(\ref{eq:local_stiffness_equation}) must be converted into the global
coordinate frame equation:
\begin{equation}
    \mathbf{f}^e=K^e\mathbf{u}^e\label{eq:global_stiffness_equation}
\end{equation}
which relates displacement to force for a single element $e$ in the global
coordinate frame.  Let $\mathbf{x}$ be a vector in the global coordinate
system and $\overline{\mathbf{x}}$ be the equivalent vector in the element
local coordinate frame.  Then the relation between the two is given by
\parencite{olver2006}:
\begin{equation}
    S\overline{\mathbf{x}}=\mathbf{x}\qquad\text{where}\qquad
    S=\left\{\mathbf{v}_1,\mathbf{v}_2,\ldots,\mathbf{v}_n\right\}
\end{equation}
With the $\mathbf{v}$ vectors forming a basis for the element centered
coordinate system.  In particular, $v_1$ is the unit vector along the element.
If the other columns of $S$ are chosen such that $S$ is orthonormal then
$S^{-1}=S^{T}$.  Therefore, the reverse transformation is:
\begin{equation}
    \overline{\mathbf{x}}=S^{-1}\mathbf{x}=S^T\mathbf{x}
\end{equation}
Noting that $\mathbf{u}$ and $\mathbf{f}$ are stacked vectors that rotate
though each coordinate twice the appropriate transformation matrix for element
$e$ becomes:
\begin{equation}
    T^e=\begin{pmatrix}
        S^T&O\\O&S^T
    \end{pmatrix}
\end{equation}
Therefore, Eq.~(\ref{eq:global_to_local}) shows the relation between the
global coordinate frame and the local coordinate frame for the force and
displacement vectors.
\begin{equation}
    \overline{\mathbf{u}}^e=T^e\mathbf{u},\qquad
    \overline{\mathbf{f}}^e=T^e\mathbf{f}
    \label{eq:global_to_local}
\end{equation}
By combining Eq.~(\ref{eq:local_stiffness_equation},
\ref{eq:global_stiffness_equation}, \ref{eq:global_to_local}) the global
stiffness matrix (for element $e$), $K^e$ can be related to the local
stiffness matrix $\overline{K}^e$ by:
\begin{equation}
    K^e=(T^e)^T\overline{K}^eT^e
\end{equation}

@@

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{listings}

@@

Most of the following formulation follows directly from
\parencite{felippa2004ch2} and \parencite{felippa2004ch3}.

@@

The direct stiffness method involves 8 steps.  The first two are purely
conceptual and thus are never really performed and the last one is a post
processing step performed after a solution is found.  These eight steps are
outlined below (taken from Figure 2.5 of \cite{felippa2004ch2}).
\begin{enumerate}[noitemsep]
    \item Disconnection
    \item Localization
    \item Element Formation
    \item Globalization
    \item Merge
    \item Application of Boundary Conditions
    \item Solution
    \item Recovery of Derived Quantities
\end{enumerate}

@@

Before the master stiffness matrix can be formed,
Eq.~(\ref{eq:local_stiffness_equation}) must be converted into the global
coordinate frame equation:
\begin{equation}
    \mathbf{f}^e=K^e\mathbf{u}^e\label{eq:global_stiffness_equation}
\end{equation}
which relates displacement to force for a single element $e$ in the global
coordinate frame.  Let $\mathbf{x}$ be a vector in the global coordinate
system and $\overline{\mathbf{x}}$ be the equivalent vector in the element
local coordinate frame.  Then the relation between the two is given by
\parencite{olver2006}:
\begin{equation}
    S\overline{\mathbf{x}}=\mathbf{x}\qquad\text{where}\qquad
    S=\left\{\mathbf{v}_1,\mathbf{v}_2,\ldots,\mathbf{v}_n\right\}
\end{equation}
With the $\mathbf{v}$ vectors forming a basis for the element centered
coordinate system.  In particular, $v_1$ is the unit vector along the element.
If the other columns of $S$ are chosen such that $S$ is orthonormal then
$S^{-1}=S^{T}$.  Therefore, the reverse transformation is:
\begin{equation}
    \overline{\mathbf{x}}=S^{-1}\mathbf{x}=S^T\mathbf{x}
\end{equation}
Noting that $\mathbf{u}$ and $\mathbf{f}$ are stacked vectors that rotate
though each coordinate twice the appropriate transformation matrix for element
$e$ becomes:
\begin{equation}
    T^e=\begin{pmatrix}
        S^T&O\\O&S^T
    \end{pmatrix}
\end{equation}
Therefore, Eq.~(\ref{eq:global_to_local}) shows the relation between the
global coordinate frame and the local coordinate frame for the force and
displacement vectors.
\begin{equation}
    \overline{\mathbf{u}}^e=T^e\mathbf{u},\qquad
    \overline{\mathbf{f}}^e=T^e\mathbf{f}
    \label{eq:global_to_local}
\end{equation}
By combining Eq.~(\ref{eq:local_stiffness_equation},
\ref{eq:global_stiffness_equation}, \ref{eq:global_to_local}) the global
stiffness matrix (for element $e$), $K^e$ can be related to the local
stiffness matrix $\overline{K}^e$ by:
\begin{equation}
    K^e=(T^e)^T\overline{K}^eT^e
\end{equation}

@@

The master stiffness equation Eq.~(\ref{eq:master_stiffness_equation}) relates
the displacement of every node, stored in $\mathbf{u}$ to every external force
in vector $\mathbf{f}$ using the linear transformation $K$.  The assembly of
each $K^e$ into $K$ is governed by \parencite{felippa2004ch3}:
\begin{enumerate}[noitemsep]
    \item Compatibility of displacements at any given node.
    \item Equilibrium of internal and external forces.
\end{enumerate}
The typical method of handling assembly of the global stiffness equations,
Eq.~(\ref{eq:global_stiffness_equation}), is to sum the forces and sum the
$K^e$ matrices by first expanding them to the size of $K$ by adding rows and
columns of zeros in the appropriate locations.  In the case of the computer
code that accompanies this document \parencite{shannon2016} the method in
section 3.5.1 of \cite{felippa2004ch3} is used.  This is know as
\emph{Assembly by Freedom Pointers}.  The algorithm is governed by
Eq.~(\ref{eq:global_to_master}) copied from \cite{felippa2004ch3} and fills $K$
by iterating through each element $e$ and then through the values in $K^e$
where $d=4$ in the 2-dimensional case and $d=6$ in the 3-dimensional case.
Let the number of elements in the truss be given by $m$.

\begin{minipage}{\textwidth}
\begin{equation*}
    K_{pq}=\sum_{e=1}^{m}K_{ij}^e
\end{equation*}
\begin{equation}
    \text{for}\label{eq:global_to_master}
\end{equation}
\begin{equation*}
    i=1,2,\ldots,d,\quad{}j=1,2,\ldots,d\qquad{}
    p=EFT^e(i),\quad{}q=EFT^e(j)
\end{equation*}
\vspace{0.1em}
\end{minipage}

The \emph{element freedom table} ($EFT^e$) is unique for each element $e$ and
gives the row/column indices where the values in $K^e$ are to be added into
$K$.  The specific algorithm to get $EFT^e$ is better explained in code and
can be found at \lstinline$libs/element_freedom_table.m$ in
\cite{shannon2016}.  In reality, $K$ is initialized to the zero matrix.  The
force vector $\mathbf{f}$ is simply constructed by summing the known (applied)
external forces in the proper indices which can also be obtained from the
\emph{element freedom table}.
